---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# labelr

The goal of labelr is to make labelling data frames easier. Think of using different labels in `ggplot2` because of different languages or because of a different audience. Or just because you want to keep things more dynamic.

## Installation

You can install it like this (not on CRAN):

``` r
devtools::install_github("tinino/labelr")
```

## How to use `labelr`

Let's take `iris`. We have `Species` as factor.
```{r}
str(iris)
```

Now assume you want to change these values, i.e. give it a new label (maybe translate to different languages or whatever). Here, we want to change labels to titlecase. Define a data frame with a column named `from` and a column named `to`:

```{r}
(species_from_to <- data.frame(
  from = levels(iris$Species), 
  to = stringr::str_to_title(levels(iris$Species))
))
```

Now set the labels. For convencience, make sure your entry is named like the variable (attention: `Species` is titlecase!):
```{r}
labelr::labels$set(Species = species_from_to)
```

Now label `iris`:
```{r}
labelr::label_df(iris)[c(1:3, 51:53, 101:103), ]
```

Now immagine you want to be able to switch labels using other distinctive parameters, like language. Here, an example might be uppercase as a second representation. To achieve this, just add another column, for example `casetype`:
```{r}
(species_from_to_titlecase <- data.frame(
  from = levels(iris$Species), 
  to = stringr::str_to_title(levels(iris$Species)),
  casetype = "titlecase"
))
# set that (hence replace previous):
labelr::labels$set(Species = species_from_to_titlecase)

# and add uppercase:
(species_from_to_uppercase <- data.frame(
  from = levels(iris$Species), 
  to = toupper(levels(iris$Species)),
  casetype = "uppercase"
))

# append this (makes it also new default):
labelr::labels$append(Species = species_from_to_uppercase)
```

Now, default is `uppercase`:
```{r}
labelr::label_df(iris)[c(1:3, 51:53, 101:103), ]
```

But we can also ask for titlecase:
```{r}
labelr::label_df(iris, casetype = "titlecase")[c(1:3, 51:53, 101:103), ]
```

Note, that the levels are still the same and will remain in the factor-order, if you specify it as a factor. If you however label a character value, it will return a factor according to the order you set in `labels$set()`. This is nice for use with, for example, `ggplot2`, where you might want to achieve a specific default order.
